Smoking and COVID Living Review (v12): a bayesian analysis

### 2021-07-27 reports/bayesian_report.Rmd

```{r load_dictionary}
source(here("scripts", "author_dictionary.R"))
```
## Download data
This is relatively automated in this script however, if you don't have an API key you can download the file individually and store it in the `/data_spreadsheet` folder as `data_extraction_current.xlsx`, the sheets are then extracted from the file.
``` {r google_data}
google_api <- rstudioapi::askForSecret("Google API Key")
register_google(google_api)
drive_download("https://docs.google.com/spreadsheets/d/1VUY5xDLBpjNBovveg9wQrc0wbLhnEkmDPMR63rfj2-Y/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)

@@

## Clean data and save updated versions as RDS files
The data from the google sheets document is then cleaned using a script from the `linelist` package which has been developed by T. Jombart and Z. Kamvar within the [R Epidemics Consortium](https://www.repidemicsconsortium.org/)
The function `clean_data()` ensures consistency among column names, country names, dates and missing values in the included sheets. Data is also saved as `.rds` files in the `/data_clean` folder of the project repository. 
```{r clean_data}
source(here("scripts", "clean_data.r"))
```
The package `flextable` can save tables in a format that is easily manageable within word which we currently use to write the manuscript (until I get better with RMarkdown). At various points I save these to the `output_data` folder.
``` {r save_as_table_for_word, include=FALSE}
datatable(a)

@@

This is a searchable table.
``` {r descriptives}
source(here("scripts", "remove_excluded.r")) # This removes the excluded studies from each table that has been read in previously.
current_total <- nrow(a)
datatable(a)

@@

save_as_docx(a, path = here('output_data', 'Countries_performing_studies.docx'))  
```
## A map is produced for where the studies are performed
``` {r interactive global map, cache = TRUE, quietly = TRUE, echo=FALSE, results = "hide"}
library(rgdal)
library(leaflet)
library(coronavirus)

my_spdf <- readOGR( 
  dsn= paste0(here::here("data_clean", "world_shape_file")), 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

from_countries <- c("USA", "UK", "Iran",  "South Korea")
to_countries <- c("United States", "United Kingdom", "Iran (Islamic Republic of)", "Korea, Republic of")
country_map <- country %>%
  mutate(Country = plyr::mapvalues(Country, from = from_countries,
                to = to_countries))

@@

``` {r produce_map}
summary_df <- coronavirus %>% 
  filter(type == "confirmed") %>%
  group_by(country) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>%
  rename("NAME" = country) %>%
  mutate(NAME = plyr::mapvalues(NAME, from = from_countries,
                to = to_countries))

my_spdf@data <- left_join(my_spdf@data, country_map %>%
            rename(NAME = Country),
          by = "NAME") %>%
  left_join(., summary_df %>%
              mutate(total_cases = total_cases/1000),
            by = "NAME") %>%
  mutate(NAME = as.factor(NAME))

my_spdf@data$Number[ which(my_spdf@data$Number == 0)] = NA
my_spdf@data$total_cases[ which(is.na(my_spdf@data$total_cases))] = 0

my_bins <- c(0, 1, 5, 10, 20, 50, 70, 110)
mypalette <- leaflet::colorBin(palette="YlOrBr", domain=my_spdf@data$Number, na.color="transparent", bins=my_bins)

mytext <- paste(
  "Country: ", my_spdf@data$NAME,"<br/>", 
  "Number of studies included: ", my_spdf@data$Number, "<br/>", 
  "COVID-19 reported cases (thousands): ", round(my_spdf@data$total_cases, 2), 
  sep="") %>%
  lapply(htmltools::HTML)

html_map <- leaflet(my_spdf) %>% 
  addTiles()  %>% 
  setView(lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(Number), 
    stroke=TRUE, 
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=mypalette, values=~Number, opacity=0.9, title = "Number of studies included", position = "bottomright" )

@@

## Classification of COVID
``` {r diagnosis_method}
diagnostics <- table(data_study_general$diagnostic_approach)
diagnostics
```

# Smoking descriptive data {.tabset .tabset-fade .tabset-pills}
``` {r smokers}
#Studies reporting smokers
current_smok <- table_1 %>%
  filter(., current_smoker != 'NA')
```
* **Current smokers:** `r paste(nrow(current_smok))` studies report on the numbers of current smokers
``` {r former smokers}
#Studies reporting former smokers
former_smok <- table_1 %>%
  filter(., former_smoker != 'NA')
```
* **Former smokers:** `r paste(nrow(former_smok))` studies report on the numbers of former smokers
```{r never smokers}
#Studies reporting never smokers
never_smok <- table_1 %>%
  filter(., never_smoker != 'NA')
```
* **Never smokers:** `r paste(nrow(never_smok))` studies report on the numbers of never smokers
``` {r current former smokers}
#Studies reporting current/former smokers
current_former_smok <- table_1 %>%
  filter(., current_former_smoker != 'NA')
```
* **Current/former smokers:** `r paste(nrow(current_former_smok))` studies report on the numbers of current/former smokers without distinguishing between them, often termed as yes to a smoking history question.
``` {r missing/never/unknown/not stated}
#Studies reporting missing data
missing_smok <- table_1 %>%
  filter(., missing != 'NA') %>%
  mutate(percent = missing/total*100)
minimum <- min(missing_smok$percent)
maximum <- max(missing_smok$percent)
#Studies reporting never/unknown
never_smok_unknown <- table_1 %>%
  filter(., never_smoker_unknown != 'NA')

@@

## Studies reporting current, former and never smoking status
``` {r complete}
#Studies reporting current, former and never smoking status
full_smoking_status <- table_1 %>%
  filter(lead_author %in% current_smok$lead_author) %>%
  filter(lead_author %in% former_smok$lead_author) %>%
  filter(lead_author %in% never_smok$lead_author)

b <- full_smoking_status %>%
  select(lead_author, sample_size, current_smoker, former_smoker, never_smoker, study_id) %>%
  rename('Lead author' = lead_author,
         'Sample size' = sample_size,
         'Current smokers' = current_smoker,
         'Former smokers' = former_smoker,
         'Never smokers' = never_smoker,
         'Study ID' = study_id)

@@

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_double(j = c("Sample size", "Current smokers", "Former smokers", "Never smokers"),
                digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

``` {r incomplete}
#Studies reporting current or current/former and never smoking
semi_full_smoking_status <- table_1 %>%
  filter(lead_author %in% current_former_smok$lead_author) %>%
  filter(lead_author %in% never_smok$lead_author) %>%
  filter(!lead_author %in% full_smoking_status$lead_author)

@@

#Smoking prevalence by country
country_prevalence_list <- table_1 %>%
    group_by(country, .add = T) %>%
  mutate(., current_smok_percentage = current_smoker/total*100) %>%
  mutate(., former_smok_percentage = former_smoker/total*100) %>%
  mutate(., missing_percentage = missing/total*100) %>%
  mutate(., not_stated_percentage = not_stated/total*100) %>%
  mutate(never_smoker_percentage = never_smoker/total*100) %>%
  group_split()

@@

## Table 1: Included studies
``` {r table_1}
#table 1 for references
table_1_refs <- table_1 %>%
  left_join(., data_study_general %>%
              select(lead_author, doi),
            by = "lead_author")
#Updating table 1
table_1_word <- table_1 %>%
  mutate(., current_percentage = current_smoker/total*100,
         former_percentage = former_smoker/total*100,
         current_former_percentage = current_former_smoker/total*100,
         never_smoker_percentage = never_smoker/total*100,
         never_smoker_unknown_percentage = never_smoker_unknown/total*100,
         not_stated_percentage = not_stated/total*100,
         missing_percentage = missing/total*100) %>%
  select(lead_author, date_published, country, sample_size, median_age, iqr_lower, iqr_upper, mean_age, lower_range,
         upper_range, standard_deviation, female_sex_percent, current_percentage, former_percentage, current_former_percentage,
         never_smoker_percentage, never_smoker_unknown_percentage, not_stated_percentage, missing_percentage, study_id) %>%
  replace_na(., list(not_stated_percentage = 0, missing_percentage = 0)) %>%
  mutate(., missing_percentage = not_stated_percentage + missing_percentage,
         missing_percentage = formatC(missing_percentage, digits = 2, format = "f"))

@@

#Data missingness
missingness <- table_6 %>%
  select(lead_author) %>%
  left_join(., table_1_word %>%
              select(lead_author, missing_percentage)) %>%
  mutate(missing_percentage = as.numeric(missing_percentage),
         low_missingness = ifelse(missing_percentage >= 20, "No", "Yes"))

#Defining poor quality
poor_quality <- missingness %>%
  select(-missing_percentage) %>%
  left_join(., smoking_status, by = "lead_author") %>%
  mutate(study_quality = ifelse(low_missingness == "No" | complete_status == "No", "Poor", "Not_poor"))

good_quality <- table_6 %>%
  left_join(., poor_quality, by = "lead_author") %>%
  select(lead_author, study_quality, biochemical_verification, random_representative_sample) %>%
  mutate(study_quality_final = ifelse(study_quality == "Not_poor" & (biochemical_verification == "Yes" | random_representative_sample == "Yes"), "good",
                                      ifelse(study_quality == "Poor", "poor", "fair")))

@@

table_1_word <- left_join(table_1_word, a, by = 'study_id') %>%
  select(1:4, 22, 5:19, 21, 20)
table_1_word$date_published <- as.Date.character(table_1_word$date_published)
write_rds(table_1_word, here::here('data_clean', 'table_1_word.rds'))

@@

a <- a %>%
  select(study_id, lead_author, date_published, country, sample_size, study_setting, median_mean, mean_used, iqr,
         female_sex_percent, current_percentage, former_percentage, current_former_percentage, never_smoker_percentage,
         never_smoker_unknown_percentage, missing_percentage, study_quality_final) %>%
  mutate(median_mean = paste(median_mean, mean_used, sep = ''),
         median_mean = na_if(median_mean, NA)) %>%
  select(-mean_used) %>%
  mutate(median_mean = ifelse(median_mean == 'NA', 'NA', paste(paste(median_mean, iqr, sep = ' ('),')', sep = ''))) %>%
  select(-iqr) %>%
  left_join(., table_1_refs %>%
              select(lead_author, doi),
            by = "lead_author") %>%
  rename('Study ID' = study_id,
         'Lead author' = lead_author,
         'Date published' = date_published,
         'Country' = country,
         'Sample size' = sample_size,
         'Study setting' = study_setting,
         'Median (IQR)' = median_mean,
         'Female %' = female_sex_percent,
         'Current smoker %' = current_percentage,
         'Former smokers %' = former_percentage,
         'Current/former smokers %' = current_former_percentage,
         'Never smokers %' = never_smoker_percentage,
         'Never/unknown smokers %' = never_smoker_unknown_percentage,
         'Missing %' = missing_percentage,
         "Study quality" = study_quality_final)

@@

numeric_columns <- c('Median (IQR)', 'Female %', 'Current smoker %','Former smokers %', 'Current/former smokers %',
                     'Never smokers %', 'Never/unknown smokers %', 'Missing %')

saveRDS(a %>%
          select(doi, everything()) %>%
          select(-"Study ID") %>%
          rename("ref" = doi),
        file = here("data_clean", "table_1_refs.rds"))
source(here("scripts", "references_in_table.r"))
```
The total number of included studies is `r paste(nrow(table_1_word))`. This table is searchable. The references are added in automatically from Zotero. The bibtex file is saved as `manuscript.bib` and is updated following each search.
```{r}
datatable(a) %>%
  formatRound(
    columns = c(8:14),
    digits = 2,
    interval = 3,
    mark = ",",
    dec.mark = getOption("OutDec")
)

@@

saveRDS(a, file = here("reports", "tables", "table_1.rds"))
save_as_docx(a, path = here('output_data', 'Table_1.docx'))
```

@@

#Table 2
table_2_word <-  table_2 %>%
  mutate(., sample = contributing_sample) %>%
  mutate(., negative_test_percentage =formatC(negative_test/sample*100, digits = 2, format = "f"),
         negative_current_percentage = formatC(negative_current_smoker/negative_test*100, digits = 2, format = "f"),
         negative_former_smoker_percentage = formatC(negative_former_smoker/negative_test*100, digits = 2, format = "f"),
         negative_current_former_smoker_percentage = formatC(negative_current_former_smoker/negative_test*100, digits = 2, format = "f"),
         negative_never_smoker_percentage = formatC(negative_never_smoker/negative_test*100, digits = 2, format = "f"),
         negative_not_stated_percentage = formatC(negative_not_stated/negative_test*100, digits = 2, format = "f"),
         positive_test_percentage = formatC(positive_test/sample*100, digits = 2, format = "f"),
         positive_current_smoker_percentage = formatC(positive_current_smoker/positive_test*100, digits = 2, format = "f"),
         positive_former_smoker_percentage = formatC(positive_former_smoker/positive_test*100, digits = 2, format = "f"),
         positive_current_former_smoker_percentage = formatC(positive_current_former_smoker/positive_test*100, digits = 2, format = "f"),
         positive_never_smoker_percentage = formatC(positive_never_smoker/positive_test*100, digits = 2, format = "f"),
         positive_not_stated_percentage = formatC(positive_not_stated/positive_test*100, digits = 2, format = "f")) %>%
  select(-data_on_testing, -missing, -date_published, -sample) %>%
  write_rds(here::here('data_clean', 'table_2_word.rds'))

@@

numeric_columns <- c('Population_tested', 'SARS_CoV_2_negative', 'N_current_smoker', 'N_former_smoker',
                     'N_current_former_smoker', 'N_never_smoker', 'N_not_stated', 'SARS_CoV_2_positive', 'P_current_smoker',
                     'P_former_smoker', 'P_current_former_smoker', 'P_never_smoker', 'P_not_stated')
a <- flextable(a) %>%
    colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
```
The number of studies reporting on infection by smoking status is: `r paste(nrow(table_2_word))`
``` {r}
a <- set_header_labels(a,
                       Population_tested = 'Total population tested',
                       SARS_CoV_2_negative = "N (%)",
                       N_current_smoker = "Current smoker (%)",
                       N_former_smoker = "Former smoker (%)",
                       N_current_former_smoker = "Current/former smoker (%)",
                       N_never_smoker = "Never smoker (%)",
                       N_not_stated = "Not stated (%)",
                       SARS_CoV_2_positive = "N (%)",
                       P_current_smoker = "Current smoker (%)",
                       P_former_smoker = "Former smoker (%)",
                       P_current_former_smoker = "Current/former smoker (%)",
                       P_never_smoker = "Never smoker (%)",
                       P_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 6, 6)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')  %>%
  colformat_char(na_str = '-')
a

@@

The ECDF shows the probability of the risk being less than or equal to RR = 0.9, `r paste(round(m1_a_ecdf(0.9)*100, 2), "%", sep = "")`

@@

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m2_a_ecdf(1.1))*100, 2), "%", sep = "")`

@@

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m3_a_ecdf(1.1))*100, 2), "%", sep = "")`

@@

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m5_a_ecdf(1.1))*100, 2), "%", sep = "")`

@@

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m7_a_ecdf(1.1))*100, 2), "%", sep = "")`

@@

The ECDF shows the probability of the risk being greater than or equal to RR = 1.1, `r paste(round((1-m8_a_ecdf(1.1))*100, 2), "%", sep = "")`

>>>>

```{r load_dictionary}
source(here("scripts", "author_dictionary.R"))
```
## Download data
This is relatively automated in this script however, if you don't have an API key you can download the file individually and store it in the `/data_spreadsheet` folder as `data_extraction_current.xlsx`, the sheets are then extracted from the file.
``` {r google_data}
drive_download("https://docs.google.com/spreadsheets/d/1fmTEW0lHYDk1OE3fw9-mJ6xu9EgBPnYoaL05-OTRiOk/edit?usp=sharing", path = here("data_spreadsheet", "data_extraction_current.xlsx"), overwrite = T)

@@

## Clean data and save updated versions as RDS files
The data from the google sheets document is then cleaned using a script from the `linelist` package which has been developed by T. Jombart and Z. Kamvar within the [R Epidemics Consortium](https://www.repidemicsconsortium.org/)
The function `clean_data()` ensures consistency among column names, country names, dates and missing values in the included sheets. Data is also saved as `.rds` files in the `/data_clean` folder of the project repository. 
```{r clean_data}
source(here("scripts", "clean_data.R"))
```
The package `flextable` can save tables in a format that is easily manageable within word which we currently use to write the manuscript (until I get better with RMarkdown). At various points I save these to the `output_data` folder.
``` {r save_as_table_for_word, include=FALSE}
datatable(a)

@@

This is a searchable table.
``` {r descriptives}
source(here("scripts", "remove_excluded.R")) # This removes the excluded studies from each table that has been read in previously.
current_total <- nrow(a)
datatable(a)

@@

save_as_docx(a, path = here('output_data', 'Countries_performing_studies.docx'))  
```
## A map is produced for where the studies are performed
``` {r interactive global map, cache = TRUE, quietly = TRUE, echo=FALSE, results = "hide"}
from_countries <- c("USA", "UK", "South Korea", "Czechia")
to_countries <- c("United States", "United Kingdom", "Korea", "Czech Rep.")
country_map <- country %>%
  mutate(Country = plyr::mapvalues(Country, from = from_countries,
                to = to_countries))

@@

``` {r produce_map}
summary_df <- coronavirus %>% 
  filter(type == "confirmed") %>%
  group_by(country) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>%
  rename("name" = country) %>%
  mutate(name = recode(name,
         "US" = "United States",
         "Czechia" = "Czech Rep.",
         "Korea, South" = "Korea"))

country_plot <- left_join(country_map  %>%
            rename("name" = Country),
            countries %>%
              select(name),
          by = "name") %>%
  left_join(., summary_df %>%
              mutate(total_cases = total_cases/1000),
            by = "name") %>%
  mutate(name = as.factor(name)) %>%
  st_as_sf()

country_plot$Number[ which(country_plot$Number == 0)] = NA
country_plot$total_cases[ which(is.na(country_plot$total_cases))] = 0

my_bins <- c(0, 1, 5, 10, 20, 50, 70, max(country$Number))
mypalette <- colorBin(palette="YlOrBr", domain = country_plot$Number, na.color="transparent", bins=my_bins)

mytext <- paste(
  "Country: ", country_plot$name, "<br/>", 
  "Number of studies included: ", country_plot$Number, "<br/>", 
  "COVID-19 reported cases (thousands): ", round(country_plot$total_cases, 2), 
  sep="") %>%
  lapply(htmltools::HTML)

html_map <- leaflet(country_plot) %>% 
  addTiles()  %>% 
  setView(lat=10, lng=0 , zoom=2) %>%
  addPolygons( 
    fillColor = ~mypalette(Number), 
    stroke=TRUE, 
    fillOpacity = 0.9, 
    color="white", 
    weight=0.3,
    label = mytext,
    labelOptions = labelOptions( 
      style = list("font-weight" = "normal", padding = "3px 8px"), 
      textsize = "13px", 
      direction = "auto"
    )
  ) %>%
  addLegend( pal=mypalette, values=~Number, opacity=0.9, title = "Number of studies included", position = "bottomright" )

@@

## Classification of COVID
``` {r diagnosis_method}
diagnostics <- table(data_study_general %>%
                       mutate(diagnostic_approach = case_when(is.na(diagnostic_approach) ~ "not_specified",
                                                              TRUE ~ diagnostic_approach)) %>%
                                pull(diagnostic_approach))
diagnostics
```

# Smoking descriptive data {.tabset .tabset-fade .tabset-pills}
``` {r smokers}
#Studies reporting smokers
current_smok <- table_1 %>%
  filter(current_smoker != 'NA' | current_vaper != "NA" | current_smoker_vaper != 'NA' | current_smoker_former_vaper != 'NA')
```
* **Current smokers:** `r paste(nrow(current_smok))` studies report on the numbers of current smokers
``` {r former smokers}
#Studies reporting former smokers
former_smok <- table_1 %>%
  filter(., former_smoker != 'NA' | former_smoker_vaper != 'NA' | former_smoker_former_vaper != 'NA')
```
* **Former smokers:** `r paste(nrow(former_smok))` studies report on the numbers of former smokers
```{r never smokers}
#Studies reporting never smokers
never_smok <- table_1 %>%
  filter(., never_smoker != 'NA' | never_smoker_vaper != 'NA' | never_smoker_former_vaper != 'NA')
```
* **Never smokers:** `r paste(nrow(never_smok))` studies report on the numbers of never smokers
``` {r current former smokers}
#Studies reporting current/former smokers
current_former_smok <- table_1 %>%
  filter(., current_former_smoker != 'NA')
```
* **Current/former smokers:** `r paste(nrow(current_former_smok))` studies report on the numbers of current/former smokers without distinguishing between them, often termed as yes to a smoking history question.
``` {r missing/never/unknown/not stated}
#Studies reporting missing data
missing_smok <- table_1 %>%
  filter(., missing != 'NA') %>%
  mutate(percent = missing/total*100)
minimum <- min(missing_smok$percent)
maximum <- max(missing_smok$percent)
#Studies reporting never/unknown
never_smok_unknown <- table_1 %>%
  filter(., never_smoker_unknown != 'NA')

@@

## Studies reporting current, former and never smoking status
``` {r complete}
#Studies reporting current, former and never smoking status
full_smoking_status <- table_1 %>%
  filter(lead_author %in% current_smok$lead_author & 
           lead_author %in% former_smok$lead_author & 
           lead_author %in% never_smok$lead_author)

b <- full_smoking_status %>%
  select(lead_author, sample_size, current_smoker, former_smoker, never_smoker, current_vaper,
         current_smoker_vaper, former_smoker_vaper, never_smoker_vaper,
         current_smoker_former_vaper, former_smoker_former_vaper, never_smoker_former_vaper,
         study_id) %>%
  rename('Lead author' = lead_author,
         'Sample size' = sample_size,
         'Current smokers' = current_smoker,
         'Current vaper' = current_vaper,
         'Former smokers' = former_smoker,
         'Never smokers' = never_smoker,
         "Current smoker, current vaper" = current_smoker_vaper, 
         "Former smoker, current vaper" = former_smoker_vaper,
         "Never smoker, current vaper" = never_smoker_vaper,
         "Current smoker, former vaper" = current_smoker_former_vaper,
         "Former smoker, former vaper" = former_smoker_former_vaper, 
         "Never smoker, former vaper" = never_smoker_former_vaper,
         'Study ID' = study_id)

@@

b <- b %>%
  select(-"Study ID") %>%
  flextable() %>%
    set_caption(caption = 'Studies reporting complete smoking status') %>%
  colformat_double(j = c("Sample size", "Current smokers", "Former smokers", "Never smokers", "Current vaper",
                         "Current smoker, current vaper", "Former smoker, current vaper", "Never smoker, current vaper",
                         "Current smoker, former vaper", "Former smoker, former vaper", "Never smoker, former vaper"),
                digits = get_flextable_defaults()$digits, big.mark = get_flextable_defaults()$big.mark, na_str = get_flextable_defaults()$na_str) %>%
  set_table_properties(width = 1, layout = 'autofit')

@@

``` {r incomplete}
#Studies reporting current or current/former and never smoking
semi_full_smoking_status <- table_1 %>%
  filter(lead_author %in% current_former_smok$lead_author &
           lead_author %in% never_smok$lead_author &
           !lead_author %in% full_smoking_status$lead_author)

@@

#Smoking prevalence by country
country_prevalence_list <- table_1 %>%
    group_by(country, .add = T) %>%
  mutate(current_smok_percentage = current_smoker/total*100,
         former_smok_percentage = former_smoker/total*100,
         missing_percentage = missing/total*100,
         not_stated_percentage = not_stated/total*100,
         never_smoker_percentage = never_smoker/total*100) %>%
  group_split()

@@

## Table 1: Included studies
``` {r table_1}
#table 1 for references
table_1_refs <- table_1 %>%
  left_join(., data_study_general %>%
              select(lead_author, doi),
            by = "lead_author")
#Updating table 1
table_1_word <- table_1 %>%
  mutate(current_percentage = current_smoker/total*100,
         former_percentage = former_smoker/total*100,
         current_former_percentage = current_former_smoker/total*100,
         never_smoker_percentage = never_smoker/total*100,
         never_smoker_unknown_percentage = never_smoker_unknown/total*100,
         current_vaper_p = current_vaper/total*100,
         current_smoker_current_vaper_p = current_smoker_vaper/total*100,
         former_smoker_current_vaper_p = former_smoker_vaper/total*100,
         never_smoker_current_vaper_p = never_smoker_vaper/total*100,
         current_smoker_former_vaper_p = current_smoker_former_vaper/total*100,
         former_smoker_former_vaper_p = former_smoker_former_vaper/total*100,
         never_smoker_former_vaper_p = never_smoker_former_vaper/total*100,
         not_stated_percentage = not_stated/total*100,
         missing_percentage = missing/total*100) %>%
  select(lead_author, date_published, country, sample_size, median_age, iqr_lower, iqr_upper, mean_age, lower_range,
         upper_range, standard_deviation, female_sex_percent, 
         current_percentage, current_vaper_p, current_smoker_current_vaper_p, current_smoker_former_vaper_p,
         former_percentage, former_smoker_current_vaper_p, former_smoker_former_vaper_p,
         current_former_percentage,
         never_smoker_percentage, never_smoker_current_vaper_p, never_smoker_former_vaper_p,
         never_smoker_unknown_percentage, not_stated_percentage, missing_percentage, study_id) %>%
  replace_na(., list(not_stated_percentage = 0, missing_percentage = 0)) %>%
  mutate(., missing_percentage = not_stated_percentage + missing_percentage,
         missing_percentage = formatC(missing_percentage, digits = 2, format = "f"))

@@

#Data missingness
missingness <- table_6 %>%
  select(lead_author) %>%
  left_join(., table_1_word %>%
              select(lead_author, missing_percentage, never_smoker_unknown_percentage)) %>%
  group_by(lead_author) %>%
  mutate(missing_percentage = as.numeric(missing_percentage),
         missing = sum(missing_percentage, never_smoker_unknown_percentage, na.rm = T),
         low_missingness = ifelse(missing >= 20, "No", "Yes")) %>%
  distinct(., .keep_all = T)

#Defining poor quality
poor_quality <- missingness %>%
  select(-missing_percentage) %>%
  left_join(., smoking_status, by = "lead_author") %>%
  mutate(study_quality = ifelse(low_missingness == "No" | complete_status == "No", "Poor", "Not_poor")) %>%
  distinct(., .keep_all = T)

good_quality <- table_6 %>%
  left_join(., poor_quality, by = "lead_author") %>%
  select(lead_author, study_quality, biochemical_verification, random_representative_sample) %>%
  mutate(study_quality_final = ifelse(study_quality == "Not_poor" & (biochemical_verification == "Yes" | random_representative_sample == "Yes"), "good",
                                      ifelse(study_quality == "Poor", "poor", "fair")))  %>%
  distinct(., .keep_all = T)

@@

table_1_word <- left_join(table_1_word, a, by = 'study_id')
table_1_word$date_published <- as.Date.character(table_1_word$date_published)
write_rds(table_1_word, here::here('data_clean', 'table_1_word.rds'))

@@

a <- a %>%
  select(study_id, lead_author, date_published, country, sample_size, study_setting, median_mean, mean_used, iqr,
         female_sex_percent,
         current_percentage, current_vaper_p, current_smoker_current_vaper_p, current_smoker_former_vaper_p,
         former_percentage, former_smoker_current_vaper_p, former_smoker_former_vaper_p,
         current_former_percentage,
         never_smoker_percentage, never_smoker_current_vaper_p, never_smoker_former_vaper_p,
         never_smoker_unknown_percentage, 
         missing_percentage, study_quality_final) %>%
  mutate(median_mean = paste(median_mean, mean_used, sep = ''),
         median_mean = na_if(median_mean, NA)) %>%
  select(-mean_used) %>%
  mutate(median_mean = ifelse(median_mean == 'NA', 'NA', paste(paste(median_mean, iqr, sep = ' ('),')', sep = ''))) %>%
  select(-iqr) %>%
  left_join(., table_1_refs %>%
              select(lead_author, doi),
            by = "lead_author") %>%
  rename('Study ID' = study_id,
         'Lead author' = lead_author,
         'Date published' = date_published,
         'Country' = country,
         'Sample size' = sample_size,
         'Study setting' = study_setting,
         'Median (IQR)' = median_mean,
         'Female %' = female_sex_percent,
         'Current smoker %' = current_percentage,
         "Current vaper %" = current_vaper_p,
         "Current smoker, current vaper %" = current_smoker_current_vaper_p,
         "Current smoker, former vaper %" = current_smoker_former_vaper_p,
         'Former smoker %' = former_percentage,
         "Former smoker, current vaper %" = former_smoker_current_vaper_p,
         "Former smoker, former vaper %" = former_smoker_former_vaper_p,
         'Current/former smokers %' = current_former_percentage,
         'Never smokers %' = never_smoker_percentage,
         "Never smoker, current vaper %" = never_smoker_current_vaper_p,
         "Never smoker, former vaper %" = never_smoker_former_vaper_p,
         'Never/unknown smokers %' = never_smoker_unknown_percentage,
         'Missing %' = missing_percentage,
         "Study quality" = study_quality_final) %>%
  distinct(., .keep_all = T)

@@

numeric_columns <- c('Median (IQR)', 'Female %', 'Current smoker %', "Current vaper %", "Current smoker, current vaper %" , "Current smoker, former vaper %",
                     'Former smoker %', "Former smoker, current vaper %", "Former smoker, former vaper %",
                     'Current/former smokers %',
                     'Never smokers %', "Never smoker, current vaper %", "Never smoker, former vaper %",
                     'Never/unknown smokers %', 'Missing %')

saveRDS(a %>%
          select(doi, everything()) %>%
          select(-"Study ID") %>%
          rename("ref" = doi),
        file = here("data_clean", "table_1_refs.rds"))

# The references in table script pulls in the doi information from the manuscript.bib file in reports
source(here("scripts", "references_in_table.R"))
```
The total number of included studies is `r paste(nrow(table_1_word))`. This table is searchable. The references are added in automatically from Zotero. The bibtex file is saved as `manuscript.bib` and is updated following each search.
```{r}
datatable(a) %>%
  formatRound(
    columns = c(8:21),
    digits = 1,
    interval = 3,
    mark = ",",
    dec.mark = getOption("OutDec")
)

@@

saveRDS(a, file = here("reports", "tables", "table_1.rds"))
save_as_docx(a, path = here('output_data', 'Table_1.docx'))

@@

#Table 2
table_2_word <-  table_2 %>%
  mutate(., sample = contributing_sample,
         negative_test_percentage =formatC(negative_test/sample*100, digits = 2, format = "f"),
         negative_current_percentage = formatC(negative_current_smoker/negative_test*100, digits = 2, format = "f"),
         negative_former_smoker_percentage = formatC(negative_former_smoker/negative_test*100, digits = 2, format = "f"),
         negative_current_former_smoker_percentage = formatC(negative_current_former_smoker/negative_test*100, digits = 2, format = "f"),
         negative_never_smoker_percentage = formatC(negative_never_smoker/negative_test*100, digits = 2, format = "f"),
         negative_current_vaper_p = formatC(negative_current_vaper/negative_test*100, digits = 2, format = "f"),
         negative_current_smoker_vaper_p = formatC(negative_current_smoker_vaper/negative_test*100, digits = 2, format = "f"),
         negative_not_stated_percentage = formatC(negative_not_stated/negative_test*100, digits = 2, format = "f"),
         positive_test_percentage = formatC(positive_test/sample*100, digits = 2, format = "f"),
         positive_current_smoker_percentage = formatC(positive_current_smoker/positive_test*100, digits = 2, format = "f"),
         positive_former_smoker_percentage = formatC(positive_former_smoker/positive_test*100, digits = 2, format = "f"),
         positive_current_former_smoker_percentage = formatC(positive_current_former_smoker/positive_test*100, digits = 2, format = "f"),
         positive_never_smoker_percentage = formatC(positive_never_smoker/positive_test*100, digits = 2, format = "f"),
         positive_never_former_percentage = formatC(positive_never_former/positive_test*100, digits = 2, format = "f"),
         positive_current_vaper_p = formatC(postitive_current_vaper/positive_test*100, digits = 2, format = "f"),
         positive_current_smoker_vaper_p = formatC(positive_current_smoker_vaper/positive_test*100, digits = 2, format = "f"),
         positive_not_stated_percentage = formatC(positive_not_stated/positive_test*100, digits = 2, format = "f")) %>%
  select(-data_on_testing, -missing, -date_published, -sample, -c(negative_current_former_vaper, negative_former_vaper, positive_current_former_vaper, positive_former_vaper)) %>%
  write_rds(here::here('data_clean', 'table_2_word.rds'))

@@

numeric_columns <- c('Population_tested',
                     "SARS_CoV_2_negative", "N_current_smoker", "N_former_smoker", "N_current_former_smoker",
                     "N_never_smoker", "N_current_vaper", "N_current_smoker_vaper", "N_not_stated",
                     "SARS_CoV_2_positive", "P_current_smoker", "P_former_smoker", "P_current_former_smoker",
                     "P_never_smoker", "P_current_vaper", "P_current_smoker_vaper", "P_not_stated")
a <- flextable(a) %>%
    colformat_double(j = numeric_columns, digits = 0, na_str = '-', big.mark = ',')
```
The number of studies reporting on infection by smoking status is: `r paste(nrow(table_2_word))`
``` {r}
a <- set_header_labels(a,
                       Population_tested = 'Total population tested',
                       SARS_CoV_2_negative = "N (%)",
                       N_current_smoker = "Current smoker (%)",
                       N_former_smoker = "Former smoker (%)",
                       N_current_former_smoker = "Current/former smoker (%)",
                       N_never_smoker = "Never smoker (%)",
                       N_current_vaper = "Current vaper (%)",
                       N_current_smoker_vaper = "Current vaper and smoker (%)",
                       N_not_stated = "Not stated (%)",
                       SARS_CoV_2_positive = "N (%)",
                       P_current_smoker = "Current smoker (%)",
                       P_former_smoker = "Former smoker (%)",
                       P_current_former_smoker = "Current/former smoker (%)",
                       P_never_smoker = "Never smoker (%)",
                       P_current_vaper = "Current vaper (%)",
                       P_current_smoker_vaper = "Current vaper and smoker (%)",
                       P_not_stated = "Not stated (%)") %>%
  add_header_row(top = TRUE, values = c("","SARS-CoV-2 negative", "SARS-CoV-2 positive" ), colwidths = c(2, 8, 8)) %>%
  theme_booktabs() %>%
  fix_border_issues() %>%
  set_table_properties(width = 1, layout = 'autofit')  %>%
  colformat_char(na_str = '-')
a

@@

@@

@@

@@

@@

@@
